---
title: "Transportation airline data"
author: "Coach Skufca"
date: "2022-09-26"
output: html_notebook
---



Experiment 1 includes summarizing, looking to match another report.


## Load Packages

```{r}

library(tidyverse)
library(here)
library(janitor)
```


## Load the BOT data

We take advantage of the `here` package to get the path to the file that we stored.

Load that data into dataframe called `df2019` using the read_csv command.   Note that you do not need to unzip the file.

**NOTE** I use `clean_names` from the janitor package to get names formatted in a consistent and useable way.

```{r}
thisfile=here("data_raw","DL_SelectFields2022.zip")

df2022=read_csv(thisfile) %>% clean_names()
```

## Subsetting to data of interest

Let's focus on flights from La Guardia (airport "LGA") and eliminate cargo flights by requiring at leat 1 passenger.

Call the resultant dataframe `df9`.

```{r}
df1=df2022 %>% filter(passengers>0,origin=="LGA" | dest=="LGA")

```

We create a barchart as we have done before:

```{r}
df1 %>% ggplot(aes(unique_carrier_name)) + geom_bar() +coord_flip()
```

Nation wide?

```{r}
df2022 %>% filter(passengers>0) %>%  ggplot(aes(unique_carrier_name)) + geom_bar() +coord_flip()
```

```{r}
df2022 %>% filter(passengers>0) %>% pull(unique_carrier_name) %>% unique()
```



### Summarize data

```{r}
df2=df2022 %>% filter(passengers>0) %>% 
  group_by(carrier_name) %>% 
  summarise(total_pass=sum(passengers),  num_flights=sum(departures_performed)) 

df2
df2 %>% ggplot(aes(reorder(carrier_name,total_pass),y=total_pass)) + geom_col() +coord_flip()
```


```{r}
df3=df2 %>% filter(total_pass>1e6)
df3 %>%   ggplot(aes(reorder(carrier_name,total_pass),y=total_pass)) + geom_col() +coord_flip()
```


```{r}
library(DataExplorer)
```

```{r}
plot_histogram(df1)
```
```{r}
library(explore)
```

```{r}
explore(df2)
```



# below here is not meaninful

** Take a look at the dataset and discover why counting rows is not going to give us a count of flights. **

The visualization we want is:

```{r}
df9 %>% ggplot(aes(month)) + geom_bar(aes(weight=departures_performed))
```
Make some observations about this plot.


### A new visualization

Can you make a boxplot where the bar height is based on the number of passengers riding each month.




### Just because you can


Here is pretty, but not so useful plot.

```{r}
df9 %>% ggplot(aes(month,fill=carrier_name)) + geom_bar(aes(weight=departures_performed))
```
## Arrivals and departures from LGA

```{r}
df10=df2019 %>% filter(passengers>0,origin=="LGA" | dest=="LGA")

df10 %>% ggplot(aes(month)) + geom_bar(aes(weight=passengers))
```

```{r}
df11=df10 %>% select(month,passengers,seats, carrier_name, dest, origin)

df12=df10 %>% select(1:5, month, contains("id") )

df13=df11 %>% mutate(percent_loading = passengers/seats*100)

df13 %>% ggplot(aes(percent_loading)) +
  geom_histogram()+facet_wrap(~carrier_name,scales="free_y")

```

### Summarize data

```{r}
df15=df2019 %>% filter(passengers>0,origin=="LGA" | dest=="LGA") %>% 
  group_by(month,carrier_name) %>% 
  summarise(total_pass=sum(passengers),  num_flights=sum(departures_performed)) %>%
  mutate(average_pass_per_flight=total_pass/num_flights)


df15 
df15 %>% ggplot(aes(x=month,y=average_pass_per_flight,color=carrier_name))+
  geom_line()
```
Insights from above plot:

* Appears to be a typical pattern (across the year) for major carriers.
* Appears to be about four sizes of flights

```{r}
df15_1=df2019 %>% filter(passengers>0,origin=="LGA" | dest=="LGA") %>% 
  group_by(month,carrier_name) %>% 
  summarise(total_seats=sum(seats),  num_flights=sum(departures_performed)) %>%
  mutate(average_seats_per_flight=total_seats/num_flights)


df15_1 
df15_1 %>% ggplot(aes(x=month,y=average_seats_per_flight,color=carrier_name))+
  geom_line()
```















```{r}
df15=df2019 %>% filter(passengers>0,origin=="LGA" | dest=="LGA") %>% 
  group_by(month) %>% mutate(percent_loading = passengers/seats*100) %>%
  summarise(loading=mean(percent_loading)) 
```

```{r}
df16=df2019 %>% filter(passengers>0,origin=="LGA" | dest=="LGA") %>% 
  group_by(month,carrier_name) %>% mutate(percent_loading = passengers/seats*100) %>%
  summarise(loading=mean(percent_loading)) 
```

```{r}
df16 %>% ggplot(aes(x=month,y=loading,color=carrier_name))+
  geom_point()+
  theme_minimal()+geom_line()+
  theme(legend.position="bottom")
```

### cargo questions - which airport

```{r}
df17=df2019 %>% filter(origin_city_name=="New York, NY" ) %>% 
  group_by(month,origin) %>% 
  summarise(freight=sum(freight,na.rm = FALSE),passengers=sum(passengers),n=n()) 
df17
```